Active Learning for SAT Solver Benchmarking

### 2022-10-11 paper/paper.tex
Revise Evaluation
Rewording and minor cutting.

Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the runtime-discretization method and the machine-learning predictor.
Thereafter, we look at the results of our active-learning approach.
Selection decisions of the aforementioned approach also reveal the importance of different problem-instance families to our model.
Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
%JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
%TF: Added a sentence for clarification. We could introduce more details in section 4.3.

@@

%JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
%TF: cut the section down a bit, renamed to "runtime prediction"
An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
Therefore, we use the runtime-prediction model that produced the best results for all further experiments.
We compare runtime predictors by looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.

@@

%To obtain the results in Table~\ref{tab:mcc},
We perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning of the runtime predictor, and \SI{20}{\%} as a test set (5-fold cross-validation).
This is repeated for each solver once.
Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
We report the average performances on the test sets.

ML classification models that we have tried include random forests, decision trees, quadratic-discriminant analysis, AdaBoost, logistic regression, $k$-nearest neighbor, multi-layer perceptrons, support vector machines, and naive Bayes.
A stacking ensemble consisting of quadratic-discriminant analysis and random forests performed best.
It produces an MCC score of \SI{0.9527}{} for time-out prediction and \SI{0.8882}{} for runtime-label prediction.
The results are significantly better than any other model that we have tried regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$.

@@

For runtime discretization, we use hierarchical clustering with $k = 3$ using a log-single-link criterion for all further experiments.
Each non-time-out runtime starts in a separate interval.
We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
As already mentioned in Section~\ref{sec:main-model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
Also, experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
%Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
%JB: "similar experiment as above" - reference unclear; what do you want to say?
%TF: Removed the sentence

\subsection{Experimental Results}
\subsubsection{Tuning our algorithm.}

Our main end-to-end experiments follow the evaluation framework that has been introduced in Section~\ref{sec:evalframe}.
Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
Evaluating a particular configuration with Algorithm~\ref{alg:eval}, returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), per selection approach (Fig.~\ref{fig:annitraincolorselection}), and per stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
%JB: as far as I understand it, the lines only compare some top-level hyperparameters, but for each of these hyperparemeters, take the best (= Pareto front) out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice" in the sense of actual, individual configurations, but rather a combination of configurations
%Colors indicate the different hyper-parameter instantiations.
%To be more specific, they represent the Pareto front of all configurations with the respective instantiations.

Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of $\delta$.
%JB: as in previous comment: consistent outperformance not guaranteed in case the lines are not individual approaches, but Pareto fronts within the compared categories (which means that different parts of the line might be related to different hyperparameter settings)
%TF: The lines are in any case no individual approaches as individual approaches are only points. So, any lines in this plot are a "combination" of configs. It's only a question of how to combine.
This is expected since selection decisions are not random.
For example, we might sample more instances of one family if it benefits discrimination of solvers.
While the partially observed PAR-2 score is skewed, the prediction model can account for this.

Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), the model-based uncertainty strategy performs best in most cases.
However, the model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
Random sampling is outperformed by the model-based uncertainty strategy in all cases showing the benefit of active learning.
By actively choosing the next experiment to run based on a model, we are more accurate in predicting the PAR-2 ranking of the solvers on the underlying dataset.
%JB: we should also describe performance of random sampling, which is visible in plot as well
%TF: added a few sentences

Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking convergence criterion performs the most consistently well.
If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.

\begin{figure}[tbp!]
  \centering
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_ranking.pgf}
    }
    \caption{
    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given ranking approaches.
    	The lines show the best performances of approaches using runtime-label-prediction ranking and partially observed PAR-2 ranking respectively.
    }
    \label{fig:annitraincolorranking}
  \end{subfigure}
  \\
  \vspace{0.2cm}
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_selection.pgf}
    }
    \caption{
    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given instance-selection approaches.
    	The lines show the best performances of approaches using uncertainty-based, information-gain-based, and random instance selection respectively.
    }
    \label{fig:annitraincolorselection}
  \end{subfigure}
  \\
  \vspace{0.2cm}
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_stopping.pgf}
    }
    \caption{
    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given stopping criteria.
    	The lines show the best performances of approaches using the ranking-convergence criterion, Wilcoxon criterion, and a fixed subset size respectively.
    }
    \label{fig:annitraincolorstopping}
  \end{subfigure}

  \caption{
  	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams showing the performance of different hyper-parameter instantiations.
  	The x-axis shows the fraction of runtime in comparison to the time needed to evaluate a solver on all instances.
  	The y-axis shows the accuracy of the predicted rank in comparison to the true PAR-2 rank.
  	Each line entails the front of Pareto-optimal configurations with the respective instantiation.
  }
  %JB: not sure if the color palette in the plots is grayscale-/color-blind-friendly; consider changing the palette (see https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=3 ) and using different line types for each color
  %TF: done
  \label{fig:e2eallsolvers}
\end{figure}

@@

\begin{figure}[tb!]
  \centering
  \resizebox{0.78\textwidth}{!}{
    \graphicspath{{../plots/}}
    \input{../plots/anni_train_optimization_goal.pgf}
  }
  \caption{
    Scatter plot showing different instantiations of our active-learning strategy for different $\delta$ on our smaller hyper-parameter-tuning dataset.
    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ that a particular instantiation needs.
    The y-axis shows the fraction of instances that a particular instantiation selects to arrive at its conclusion.
    The color indicates the weighting between different optimization goals $\delta \in \left[0, 1\right]$.
    A small $\delta$ refers to favoring runtime over accuracy and a big $\delta$ refers to favoring accuracy over runtime.
  }
  \label{fig:annitrainoptgoal}
\end{figure}

Figure~\ref{fig:annitrainoptgoal} also shows an interesting difference between different optimization goals.
It shows how different configurations select instances with different difficulties.
If it is, on the one hand, desired to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), corresponding approaches favor selecting many \emph{easy} problem instances (fraction of instances bigger than the fraction of runtime within $\tilde{\mathcal{I}}$).
By having many observations, it is easier to build a model.
If it is, on the other hand, desired to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high $\delta$), corresponding approaches favor selecting fewer \emph{difficult} problem instances (fraction of instances smaller than the fraction of runtime within $\tilde{\mathcal{I}}$).

@@

Having selected the most promising hyper-parameters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
The best-performing approach for $\delta = 0.85$ uses runtime-label prediction ranking, model-based uncertainty sampling, and a ranking-convergence stopping criterion.
It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the runtime that would be needed to evaluate the solver on the full dataset ($O_{\operatorname{rt}}$).
This trade-off provides a better way of prototyping and evaluating new SAT solvers rather than evaluating them on a manually picked subset based on domain knowledge.
The sampled instances account for about \SI{5.24}{\%} of instances indicating the preference for selecting more difficult instances as already discussed in the previous section.
The average Spearman correlation between the PAR-2 ranking and the predicted ranking scores is about \SI{0.9572}{} indicating that the differences in PAR-2 scores are quite similar to $s_3$ scores.
Because of that, the runtime-prediction model helps determine on what problem instances solvers perform similarly or are better/worse.
%JB: many numbers, but some qualitative (and positively sounding, i.e., saying that our approach indeed is useful) interpretation would be nice
%TF: revised this section

@@

Selection decisions of our approach also reveal the importance of different problem-instance families to our model.
Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.

Figure~\ref{fig:annifinalfamilies} shows a scatter plot of problem-instance families regarding their occurrence within the dataset and the average occurrence in benchmarks created by active learning.
We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track benchmark set~\cite{sat2022}.
While most families appear by the same fraction in the dataset and in active-learning-selected benchmarks, there are a few outliers that need some further discussion.
Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our model in distinguishing solvers.
Instances of these families are selected over-proportionally in comparison to their fraction in the dataset.
In contrast to that, instances of the biggest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and in active-learning-selected benchmarks.
Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their huge weight in the dataset suggests.
A possible explanation is that \emph{cryptography} instances are too similar, which means that a smaller fraction of them is sufficient to determine a solver's performance on all of them.

@@

\begin{figure}[tb!]
  \centering
  \resizebox{0.85\textwidth}{!}{
    \graphicspath{{../plots/}}
    \input{../plots/anni_final_families.pgf}
  }
  \caption{
    Scatter plot showing the \emph{importance} of different instance families to our model.
    The x-axis shows the fraction of instances within a particular family in the complete SAT Competition~2022 Anniversary Track benchmark set~\cite{sat2022}.
    The y-axis shows the average fraction of instances within a particular family that is selected by our active-learning strategies.
    The dashed line represents families that are represented with the same fraction in, both, our dataset and the incremental benchmark sets $\tilde{\mathcal{I}}$ that are chosen by our approach.
  }
  \label{fig:annifinalfamilies}
\end{figure}

@@

In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
Given a new solver, we want to find its ranking amidst its competitors.
Our approaches provide accurate ranking predictions while only needing a fraction of the runtime resources that a full evaluation on all benchmarking instances would need.
We can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy, while only needing \SI{10}{\%} of the full-evaluation time.
We use a runtime discretization technique as this enables us to transform the regression problem of directly predicting runtimes into the much simpler notion of classification.
We have shown that, albeit being more simple, the classification of discrete runtime labels produces good results.
We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
%A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results in most cases.
%JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
%TF: commented out
We also took a brief look at which problem-instance families are the most prevalent in selection decisions.

@@

On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are agnostic to the underlying problem.
Solely the runtime-prediction model might need some tweaking.
%JB: is there some methodological generalization needed or can't we just apply the same framework (instantiating the instance-feature part differently) to these use cases?
%TF: clarified
Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires expert knowledge to hand-select instances of interest.

>>>>

Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the machine-learning predictor and the runtime-discretization method.
Thereafter, we look at the results of our active-learning approach.
Finally, we also analyze the importance of different instance families to our framework.
Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
%JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
%TF: Added a sentence for clarification. We could introduce more details in section 4.3.
%JB: sounds like a good idea

@@

%JB: we should probably remove this section here and jump right into active learning;
% - results of classifier comparison can shortly be described in Section 4 (new subsection) or 3.2 ("Solver Model", which currently mainly describes runtime discretization); in any case, classifier should be discussed earlier
% - in current version, we already state in Section 3.2 that we use hierarchical clustering, so description of algorithm and some "performed well in preliminary experiments" could go there
An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
Therefore, we compare predictors first and select one runtime-prediction model for all further experiments.
We evaluate predictors with the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.

@@

%To obtain the results in Table~\ref{tab:mcc},
We perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning of the runtime predictor, and \SI{20}{\%} as a test set (5-fold cross-validation).
This is repeated for each solver once.
Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
%JB: we should state (were we introduce prediction model, i.e., either in 3.2 or a new subsection in 4) which features we use; I assume instance features + runtimes of other solvers? (else CV among solvers would not make sense for this experiment)
We compare the average performances on the test sets.

ML classification models that we have tried include random forests, decision trees, quadratic-discriminant analysis, AdaBoost, logistic regression, $k$-nearest neighbor, multi-layer perceptrons, support vector machines, and naive Bayes.
A stacking ensemble consisting of quadratic-discriminant analysis and random forests performed best.
It produces an MCC score of \SI{0.8882}{} for predicting three-class runtime labels.
% \SI{0.9527}{} for time-out prediction
The results are significantly better than any other model that we have tried regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$.

@@

For runtime discretization, we use hierarchical clustering with $k = 3$ and a log-single-link criterion for all further experiments.
Each non-time-out runtime starts in a separate interval.
We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
As already mentioned in Section~\ref{sec:main-model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
Also, experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
%Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
%JB: "similar experiment as above" - reference unclear; what do you want to say?
%TF: Removed the sentence

\subsection{Active-Learning Results}

\subsubsection{Tuning our algorithm.}

Our end-to-end experiments follow the evaluation framework introduced in Section~\ref{sec:evalframe}.
Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
%Colors indicate the different hyper-parameter instantiations.
%To be more specific, they represent the Pareto front of all configurations with the respective instantiations.

Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the runtime-accuracy trade-off parameter~$\delta$.
%JB: as in previous comment: consistent outperformance not guaranteed in case the lines are not individual approaches, but Pareto fronts within the compared categories (which means that different parts of the line might be related to different hyperparameter settings)
%TF: The lines are in any case no individual approaches as individual approaches are only points. So, any lines in this plot are a "combination" of configs. It's only a question of how to combine.
%JB: I always made the implicit assumption that the plots are based on learning curves, showing the progress within individual AL runs, and then multiple of these curves were combined (per top-level configuration); so the curves actually draw a line between different final results of AL runs? should be clarified somewhere
This is expected since selection decisions are not random.
For example, we might sample more instances of one family if it benefits discrimination of solvers.
While the partially observed PAR-2 score is skewed, the prediction model can account for this.

Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), the model-based uncertainty strategy performs best in most cases.
%JB: maybe we can drop "model-based" and use "uncertainty sampling" instead? I adapted this in Section 3, but section 4 + 5 + the plots would have to be adapted as well (else, we should revert my changes in Section 3)
However, the model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
Random sampling performs quite well but is outperformed by the model-based uncertainty strategy in all cases, showing the benefit of actively selecting instances based on a prediction model.

Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking convergence criterion performs the most consistently well.
If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
%JB: plot says "Wilcoxon" without stopping; only minor point, but we should make naming of the components of our framework consistent

\begin{figure}[tbp!]
  \centering
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_ranking.pgf}
    }
    \caption{Ranking approaches}
    \label{fig:annitraincolorranking}
  \end{subfigure}
  \\
  \vspace{0.2cm}
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_selection.pgf}
    }
    \caption{Selection approaches}
    \label{fig:annitraincolorselection}
  \end{subfigure}
  \\
  \vspace{0.2cm}
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_stopping.pgf}
    }
    \caption{Stopping criteria}
    \label{fig:annitraincolorstopping}
  \end{subfigure}

  \caption{
  	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams comparing different hyper-parameter instantiations of our active-learning framework.
  	The x-axis shows the ratio of total solver runtime on the sampled instances relative to all instances.
  	The y-axis shows the ranking accuracy (cf. Section~\ref{sec:goal}).
  	%JB: we should use "ranking accuracy" in plots, since we explicitly defined this metric earlier
  	Each line entails the front of Pareto-optimal configurations for the respective hyper-parameter instantiation.
  }
  \label{fig:e2eallsolvers}
\end{figure}

@@

\begin{figure}[tb!]
  \centering
  \resizebox{0.78\textwidth}{!}{
    \graphicspath{{../plots/}}
    \input{../plots/anni_train_optimization_goal.pgf}
  }
  \caption{
    Scatter plot comparing different instantiations of $\delta$ for our active-learning strategy on the hyper-parameter-tuning dataset.
    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axis shows the fraction of instances sampled.
    The color indicates the weighting between different optimization goals $\delta \in \left[0, 1\right]$.
    The larger $\delta$, the more we favor accuracy over runtime.
  }
  \label{fig:annitrainoptgoal}
\end{figure}

Figure~\ref{fig:annitrainoptgoal} shows an interesting consequence of weighting our optimization goals:
If we, on the one hand, desire to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), approaches favor selecting many \emph{easy} instances.
In particular, the fraction of sampled instances is bigger than the fraction of runtime.
By having many observations, it is easier to build a model.
If we, on the other hand, desire to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high $\delta$), approaches favor selecting fewer \emph{difficult} instances.
In particular, the fraction of instances is smaller than the fraction of runtime.

@@

Having selected the most promising hyper-parame\-ters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
The best-performing approach for $\delta = 0.85$ uses runtime-label prediction ranking, model-based uncertainty sampling, and a ranking-convergence stopping criterion.
%JB: we need to motivate choosing this \delta
It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the time for running the solver on the full dataset ($O_{\operatorname{rt}}$).
%JB: could we (with reasonable effort) integrate the learning curve of this particular approach as a plot?
This trade-off provides a better way of prototyping and evaluating new SAT solvers rather than evaluating them on manually picked instance based on domain knowledge.
The sampled instances account for about \SI{5.24}{\%} of instances, indicating a preference for selecting difficult instances.
The average Spearman correlation between the PAR-2 ranking and the predicted ranking is about \SI{0.9572}{}, indicating a high similarity of these two rankings.
Because of that, the runtime-prediction model helps determine on what problem instances solvers perform similarly or are better/worse.
%JB: conclusion not straightforward from prior text; even if overall estimated score (which is an average over many instances) is good, estimate on individual instances might be quite erroneous

@@

Selection decisions of our approach also reveal the importance of different instance families to our framework.
Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
While most families appear with the same fraction in the dataset and the sampled benchmarks, there are a few outliers that need further discussion.
Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
Instances of these families are selected over-proportionally in comparison to the full dataset.
In contrast to that, instances of the biggest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their huge weight in the dataset suggests.
A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.

@@

\begin{figure}[tb!]
	\centering
	\resizebox{0.85\textwidth}{!}{
		\graphicspath{{../plots/}}
		\input{../plots/anni_final_families.pgf}
	}
	\caption{
		Scatter plot showing the \emph{importance} of different instance families to our framework.
		The x-axis shows the frequency of instance families in the full Anniversary Track dataset.
		The y-axis shows the average frequency of instance families in the samples selected by active learning.
		The dashed line represents families that are occur with the same frequency in both, dataset and samples.
	}
	\label{fig:annifinalfamilies}
\end{figure}

@@

In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
Given a new solver, we want to find its ranking amidst its competitors.
Our approaches provide accurate ranking predictions while only needing a fraction of the runtime resources that a full evaluation on all benchmarking instances would need.
We can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy, while only needing \SI{10}{\%} of the full-evaluation time.
We use a runtime discretization technique as this enables us to transform the regression problem of directly predicting runtimes into the much simpler notion of classification.
We have shown that, albeit being more simple, the classification of discrete runtime labels produces good results.
We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
%A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results in most cases.
%JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
%TF: commented out
We also took a brief look at which instance families are the most prevalent in selection decisions.

@@

On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are agnostic to the underlying problem.
Solely the runtime-prediction model might need some tweaking.
Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires expert knowledge to hand-select instances of interest.
